---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rfBLT: Random Feature Bayesian Lasso Takens

<!-- badges: start -->
<!-- badges: end -->

An R package for implementing Regularized or Sparse Random Feature Models integrating with Time Delay Embedding Theorem, within both Frequentist and Bayesian frameworks. Currently, only Gaussian error models are supported. In the Frequentist setting, lasso, ridge, and elastic net regularization are available, while the Bayesian setting provides shrinkage priors for lasso and ridge. Recursive prediction is implemented for time series forecasting. The package also supports cross-sectional data and time series prediction using lagged features, without Takens’ theorem, through standard Sparse Random Feature Models.

## Installation

You can install the development version of rfBLT from [GitHub](https://github.com/thuthiminhnguyen/rfBLT) with:

``` r
# install.packages("devtools")
devtools::install_github("thuthiminhnguyen/rfBLT")
```

The package now can be loaded and used:

```{r}
library(rfBLT)
```

## Example

The following example illustrates how to fit and forecast time series values using the Sparse Random Feature Model using Takens' Theorem, within both Bayesian and Frequentist approaches.

```{r example, eval=FALSE}
# Generate data
set.seed(456)
y = cumsum(rnorm(1000))

# Visualize time series
plot(y, type="l", main="Time series data", xlab="Time index", ylab="Value")
```

### Bayesian 

The `ts_forecast_bayes_reg_rfm_taken` function fits a Bayesian model to a time series and forecasts future values for a specified prediction horizon and confidence level. The example below demonstrates the model with a smoothing method applied:

```{r, eval=FALSE}
fit_bayes_reg_rfm_taken <- ts_forecast_bayes_reg_rfm_taken(ts_data=y, pred_size=7, CI=95, smooth_diff = TRUE)
```

The element `ts_forecast_bayes_reg_rfm_taken$fit_results` returns an object of class bayesreg, which can be summarized as follows: 

```{r, eval=FALSE}
summary(fit_bayes_reg_rfm_taken$fit_results)
```

The output includes both smoothed and non-smoothed time derivatives for diagnostic purposes.

```{r, eval=FALSE}
# Plot the smoothed and non-smoothed time derivatives
plot(fit_bayes_reg_rfm_taken$raw_diff, type="l", 
     main="Smoothed and non-smoothed time derivatives",
     ylab="Time derivatives", xlab="Time index")
lines(fit_bayes_reg_rfm_taken$smooth_diff, col="blue", lwd=2)
legend("topleft",                
       legend = c("Non-smoothed", "Smoothed"),  
       col = c("black", "blue"), 
       lty = 1, lwd = c(1, 2), bty = "n") 
```

A plot function is implemented to plot Effective Sample Size `plot.ess_plot()`. 

```{r, eval=FALSE}
plot(fit_bayes_reg_rfm_taken$ess)
```

The mean predicted values obtained from the model are

```{r, eval=FALSE}
fit_bayes_reg_rfm_taken$y_pred
```

The 95% credible intervals for the forecasts are:

```{r, eval=FALSE}
fit_bayes_reg_rfm_taken$pred.ci
```

### Frequentist

The following code demonstrates a Frequentist Sparse Random Feature Model with time-delay embedding, without bootstrapped confidence intervals:

```{r, eval=FALSE}
fit_freq_reg_rfm_taken <- ts_forecast_freq_reg_rfm_taken(ts_data=y, pred_size=7, CI=95, smooth_diff = TRUE)
```

The element `fit_freq_reg_rfm_taken$fit_results` returns an object of class glmnet, and its coefficients can be summarized as follows:

```{r, eval=FALSE}
coef(fit_freq_reg_rfm_taken$fit_results)
```

Similar to the Bayesian approach, both smoothed and non-smoothed time derivatives are computed and can be visualized.

```{r, eval=FALSE}
# Plot the smoothed and non-smoothed time derivatives
plot(fit_freq_reg_rfm_taken$raw_diff, type="l", 
     main="Smoothed and non-smoothed time derivatives",
     ylab="Time derivatives", xlab="Time index")
lines(fit_freq_reg_rfm_taken$smooth_diff, col="blue", lwd=2)
legend("topleft",                
       legend = c("Non-smoothed", "Smoothed"),  
       col = c("black", "blue"), 
       lty = 1, lwd = c(1, 2), bty = "n") 
```

The prediction values and their confidence intervals (if applicable) are:

```{r, eval=FALSE}
# Predictions
fit_freq_reg_rfm_taken$y_pred

# Confidence interval
fit_freq_reg_rfm_taken$pred.ci
```

## References

```{r citation_text, results='asis', echo=FALSE}
format(citation("rfBLT"), style = "text")
```
